Enhancing diagnostic of stochastic mortality models leveraging contrast trees. An application on Italian data

### 2022-12-25 Manuscript Contrast Trees_R2.tex
ok

%-------------------------------------------------------------
\section*{Discussion}
%-------------------------------------------------------------
%\textit{The Discussion should be succinct and must not contain subheadings.}\\
\textcolor{blue}{Evaluating, and thus eventually improving, the fit of mortality models is crucial for both demographers and actuaries. 
Indeed, in particular situations, common in actuarial practice, data quality can turn the mortality estimate difficult. 
A prime example is the case of small subpopulations where a common method such as the Lee-Carter may not guarantee reliable estimation. In this sense, our proposal fills the gap between mortality modeling and model diagnostics, particularly for nontraditional modeling as a machine learning framework.\\
Contrast trees consist of a general method based on machine learning that can be applied to any model, expressed as a regression model, to evaluate the goodness of fit and identify the worst-performing regions in the input space. The main characteristic that discriminates this method from traditional diagnostic tools is automatically identifying the regions in which a given model produces a high error for certain combinations of ages and calendar years.
Well-known diagnostic tools often used in the literature to assess the goodness-of-fit of a mortality model, such as BIC and AIC, require the likelihood function, which is not available for machine learning models. Therefore, Contrast trees provide a unified approach for assessing and comparing the accuracy of traditional mortality models with machine learning algorithms.}\\
In Contrast trees, the detection of the regions in which a model worst performs can be considered an evolution of the standard analysis on residuals, in which the detection of the highest residuals is typically assigned to graphical analyzes using heatmaps and scatter plots \cite{CBD09,Villegas2018}, and to summary measures like RMSE and MAPE calculated on the overall input space and not by region.
Conversely, the decision tree structure of Contrast trees enables quantifying the discrepancy between the estimates provided by a model and the actual observations in each region identified by Contrast trees.

>>>>

%-------------------------------------------------------------
\section*{Discussion}
%-------------------------------------------------------------
%\textit{The Discussion should be succinct and must not contain subheadings.}\\
\textcolor{blue}{Evaluating, and thus eventually improving, the fit of mortality models is crucial for both demographers and actuaries. 
Indeed, in particular situations, common in actuarial practice, data quality can turn the mortality estimate difficult. 
A prime example is the case of small subpopulations where a common method such as the Lee-Carter may not guarantee reliable estimation. 
\textcolor{green}{In mortality modeling, the objective of diagnostic checking is to ascertain whether the model fits the historical data by obeying an underlying probabilistic hypothesis.  This procedure is carried out using residuals diagnosis checking with a Gaussian or more often a Poisson assumption (see, e.g., \cite{RH}).
Leveraging \cite{Friedman2020}, who introduces contrast trees to estimate the full conditional probability distribution without any parametric assumptions, we propose a prominent alternative, with particular regard to the intersection of Machine Learning and Mortality modeling fields.
In this sense, our proposal fills the gap between mortality modeling and model diagnostics, particularly for nontraditional modeling as a machine learning framework.}\\
Contrast trees consist of a general method based on machine learning that can be applied to any model, expressed as a regression model, to evaluate the goodness of fit and identify the worst-performing regions in the input space. The main characteristic that discriminates this method from traditional diagnostic tools is automatically identifying the regions in which a given model produces a high error for certain combinations of ages and calendar years.
Well-known diagnostic tools often used in the literature to assess the goodness-of-fit of a mortality model, such as BIC and AIC, require the likelihood function, which is not available for machine learning models. Therefore, Contrast trees provide a unified approach for assessing and comparing the accuracy of traditional mortality models with machine learning algorithms.}\\
In Contrast trees, the detection of the regions in which a model worst performs can be considered an evolution of the standard analysis on residuals, in which the detection of the highest residuals is typically assigned to graphical analyzes using heatmaps and scatter plots \cite{CBD09,Villegas2018}, and to summary measures like RMSE and MAPE calculated on the overall input space and not by region.
Conversely, the decision tree structure of Contrast trees enables quantifying the discrepancy between the estimates provided by a model and the actual observations in each region identified by Contrast trees.
